Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2019-04-08 man/ms.Rmd
update resilience results

## Resilience of greenness and radial tree-growth to drought events

The strength of the response of the resilience indices was related to drought severity. The exploration of resilience metrics of tree-growth for all severe droughts in our climatic data `REVIEW`(1900 - 2017), revealed a positive relation between resilience (*Rs*) and drought severity ($r^2$ = 0.336) (Figure 6). A similar pattern was found for recovery (*Rc*; $r^2$ = 0.438) but negative for resistance (*Rt*, $r^2$ = 0.155). Resilience values of tree-growth for 2005 was the lowest of the drought events analyzed (Figure 6) even not being the most severe drought events. `REVIEW` (CONEXION con el siguiente parrafo)

*Q. pyrenaica* forest showed significantly lower resistance values (*Rt*) to the 2005 drought event than to that in 2012 for greenness and for radial growth (Table 3; Figure 7). The 2005 drought reduced greenees and growth more than that of 2012 (Tables S1 and S2). Resistance values to drought for greenness and tree-growth varied between sites (Table 3). Southern populations showed significantly higher values of resistance than northern ones (Tables S1 and S2). It was particularly important for the 2005 drought event where the growth was reduced to 44.5 % respect to that of the preceding period (Figure 7).

Recovery (*Rc*) of greenness and growth were significantly different bewteen drought events and sites (Table 3). In the 3-year period after the 2005 drought, greenness was 112 % (Rc = 1.12) and after 2012 was 105.7 % (Rc = 1.057) (Table S1). An opposite pattern was found for tree-growth, with significantly lower values of recovery after the 2005 drought, staying at levels of *Rc* < 1 (Figure 7, Table S2). Northern populations showed significantly higher values of recovery than southern sites for greenness and tree-growth (Table S2). For southern populations, no significant differences were found for recovery of greenness after the 2005 and 2012 drought event (p = 0.2453; Figure 7; Table S1). Recovery values for tree-growth of southern populations were below or close to 1 (Figure 7, Table S2).

The exploration of resilience metrics of tree-growth for all severe droughts in our climatic data (period 1900-2017?),

@@

@@

>>>>

## Resilience of greenness and radial tree-growth to drought events
At long-term perspective, the strength of the response of the resilience indices was related to drought severity. The exploration of resilience metrics of tree-growth for all severe droughts in our climatic data (1900 - 2017) (Table S3), revealed a positive relation between resilience (*Rs*) and drought severity ($r^2$ = 0.336, p = 0.048) (Figure 6). A similar pattern was found for recovery (*Rc*; $r^2$ = 0.438, p = 0.019) but negative for resistance (*Rt*, $r^2$ = 0.155, p = 0.205).

For greenness and tree-growth, resilience metrics were significantly different bewteen drought events at short-term (Table 3). We obtained significantly higher Resilience (*Rs*) and Resistance (*Rt*) values for the 2012, the most severe event, than for 2005 in both variables (Tables S1-S2; Figure 7). Recovery (*Rc*) showed contrasting pattern for EVI and tree-growth. Recovery values for greenness was higher for 2005 than for 2012 drought event (Table S1, Figure 7).

Resilience metrics significantly varied between sites, except for Resilience (*Rs*) of tree-growth (p = 0.534; Table 3). Southern populations showed lower values for Recovery (*Rc*) than northern site for greenness and tree-growth (Tables S1 and S2). In opposite, Resistance and Resilience values were higher for southern sites than for northern ones (Tables S1 and S2).

@@

YA ESTOY DESESPERADO CON ESTE PAPER
*Q. pyrenaica* forest showed significantly lower resistance values (*Rt*) to the 2005 drought event than to that in 2012 for greenness and for radial growth (Table 3; Figure 7). The 2005 drought reduced greenees and growth more than that of 2012 (Tables S1 and S2). Resistance values to drought for greenness and tree-growth varied between sites (Table 3). Southern populations showed significantly higher values of resistance than northern ones (Tables S1 and S2). It was particularly important for the 2005 drought event where the growth was reduced to 44.5 % respect to that of the preceding period (Figure 7).

Recovery (*Rc*) of greenness and growth were significantly different bewteen drought events and sites (Table 3). In the 3-year period after the 2005 drought, greenness was 112 % (Rc = 1.12) and after 2012 was 105.7 % (Rc = 1.057) (Table S1). An opposite pattern was found for tree-growth, with significantly lower values of recovery after the 2005 drought, staying at levels of *Rc* < 1 (Figure 7, Table S2).

Northern populations showed significantly higher values of recovery than southern sites for greenness and tree-growth (Table S2). For southern populations, no significant differences were found for recovery of greenness after the 2005 and 2012 drought event (p = 0.2453; Figure 7; Table S1). Recovery values for tree-growth of southern populations were below or close to 1 (Figure 7, Table S2).

Similar to the long-term observed pattern,

was greater after more severe drought event (2012). Opposite pattern was observed for greenness with

Lower resistance values (*Rt*) were observed to the 2005 drought event than to that in 2012 for both variables. The 2005 drought reduced greenees and growth more than that of 2012 (Tables S1 and S2). Resistance values to drought for greenness and tree-growth varied between sites (Table 3). Southern populations showed significantly higher values of resistance than northern ones (Tables S1 and S2). It was particularly important for the 2005 drought event where the growth was reduced to 44.5 % respect to that of the preceding period (Figure 7).

@@

---> Nota

Para la discussion: 
- Resilience values of RW for 2005 was the lowest of the drought events analized even not being the moste severe drought event 
- 2005 reduced greenness and bai more than 2012
